---
title: An Introduction to GitHub and R Markdown
author: Hemanth Bharatha Chakravarthy
date: '2019-03-22'
slug: introduction-to-github-and-r-markdown
categories:
  - R
tags:
  - R Markdown
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE)
```

# 1 GitHub

### 1.1 What is GitHub?

- Cloud-based version control system
- But what is version control? 
        - Simple analogy: Google Drive for developers
        - Track and manage changes to your code
        - Version control lets developers safely work through branching and merging
        - Work with online Git repository
        - Make changes in your own branch
        - Merge with online cloud branch
        - Changes are tracked and can be reverted if needed

### 1.2 Why GitHub?

- Don't really have a choice---industry standard!
- Cloud backup
- Collaboration, version control
- Don't have to reinvent the wheel: use existing Git repos
- Community review of your code
- Other uses: GitHub forum, hosting websites, GitHub Classroom, etc.

### 1.3 Setup for RStudio

**Make your account and install into RStudio**

- Sign up at www.github.com 
- In console: `install.packages("devtools")`
      - If this doesn't work, you might need to update your R
      - Update at https://www.r-project.org/
      
**Start your first project**

- Create a new GitHub Repository 
- (If collaborative) Open options in your repo -> Share with collaborators
- In the GitHub repo: Clone or Download -> Copy the repo link
- In RStudio: File -> New Project -> Version Control -> Git -> Paste the link
- Check out Git tab in top right corner alongside Environment, History, and Connections

**Quick note**

- Gitignore: what local files or filetypes should Git ignore in prompting you to Commit in the Git tab? (e.g., `*.Rproj` in your `.gitignore`)
- Readme: let people know what cool thing you're building in your `readme.md`

More information: https://www.datacamp.com/community/tutorials/git-setup

### 1.4 Three Main Git Functions

1. Commit: create a new commit containing the current contents of the index and the given log message describing the changes (https://git-scm.com/docs/git-commit)
2. Push: push the local version of directory into your Git directory, i.e, your repo (https://git-scm.com/docs/git-push)
3. Pull: pull the current online version of the project into your local directory. Used when the online version has been edited since you last clones/pulled. Maintains both the changes in the online version and your local version

Very simple guide on the workflow: http://rogerdudler.github.io/git-guide/

A more detailed guide: https://happygitwithr.com/rstudio-git-github.html

### 1.5 Working with Git from the Shell

Use `man git` in Shell for help

- Commit: `git commit -m "Commit message"`
- Push: `git push origin master`
- Pull: `git pull`

### 1.6 More Resources

- GitHub Cheat Sheet: https://education.github.com/git-cheat-sheet-education.pdf
- A solid introduction: https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1004668

# 2 R Markdown (aka Rmd)

### 2.1 What is Rmd?

"R Markdown is a file format for making dynamic documents with R. An R Markdown document is written in markdown (an easy-to-write plain text format) and contains chunks of embedded R code, like the document below." - https://rmarkdown.rstudio.com/articles_intro.html

### 2.2 Why Rmd?

- This is written in Rmd! So cool!
- Using Rscripts is cumbersome: Rmd lets you analyze and present data in one go
- Can publish as html, pdf, and word doc
- Automatically LaTeXs if you publish to pdf
- Can handle complex scientific and mathematic notation

### 2.3 Setup

Very simple. Next time you make a new file, make it a R Markdown file instead of a R script and save it as `x.Rmd`

### 2.4 Chunks and Chunk Options

**Chunks**

A mini-R-script in the middle of your text, headings, etc. in Markdown

E.g.:

```{r demo_markdown, include=TRUE, message=FALSE}
library(tidyverse)
# Making a random dummy dataframe
x <- data.frame(replicate(5,sample(0:1,1000,rep=TRUE)))
# Print three rows of our dummy df
x[1:3,]
```

**Chunk Options**

You might not want to show your code, the warnings, and messages but show the output--Rmd let's you control that.

E.g.: There's a hidden chunk below but you'll only see the table it outputs

```{r echo=FALSE, warning=FALSE, message=FALSE}
options(tinytex.verbose = TRUE)
library(gt)
library(dplyr)

x %>% 
  group_by(X1) %>% 
  summarise() %>% 
  gt() %>% 
  tab_header("Cool table")
```


**Global Options**

You can also initialize a setup chunk with global chunk options like below.

`knitr::opts_chunk$set(include = TRUE, warning = FALSE) # Random chunk options`

Here's a [cheat sheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)!

### 2.5 Knitting to Different File Formats

- `Cmd or Ctrl + Shift + K` to knit to chosen format
- Use the Knit button in the RStudio toolbar to knit to some other format
- Alternately, edit the YAML header at the top of the file. Options include:

    `output:`

        pdf_document: default
  
        word_document: default
  
        html_document: default

### 2.6 More Resources

- Xie, Yihue. [R Markdown: The Definite Guide](https://bookdown.org/yihui/rmarkdown/yihui-xie.html) (written in [Bookdown](https://bookdown.org/yihui/bookdown/)!)
- R for Data Science (R4DS) [Chapter](https://r4ds.had.co.nz/r-markdown.html) on Rmds